### 2019-07-08 papers/journal/encoding.tex
Update on the fig 2 legend
Fixed missing part objection

%if all the pairs of the present molecule are present on the same edge than all of them inhibit each other resulting in making all the molecules inactive, we term this process ``pairing inhibition".
%
The constraints for the variants can be given to a SAT or a SMT solver to find
VTSs that belong to the variants.
\subsection{Encoding for $\fqbf$ search problem}\label{enc:univ-cond}
%
\input{table2}
\noindent The specification for universal condition Section~\ref{sec:property} requires quantifier alternation. 
%
Most SAT solvers can not efficiently handle the quantifier in the formulas. 
% 
Although, it is possible to encode the universal condition as a SAT or SMT problem.
%
There is always a danger of a blow-up in the formula size, making it harder for the SAT solver to handle. 
%Which may blow up the formula size for the solver to solve.
% 
As an experiment, we have implemented a direct SMT version (encoding not presented) of the universal problem extending the current $\fsmt$ search encodings.
%
The basic idea is to conjunct the formula for each combination value of the quantified variable. 
%
The Table~\ref{tab:satqbf-graph}, shows the search results of this encoding for the universal condition with 4 nodes and 3-connected graphs. 
%
The connectivity constraint and connectedness constraints are tagged with subscript $U$, $\texttt{StabilityC}_{U}$, $\texttt{Connected}(k)_{U}$,  to distinguish from earlier presented encodings of existential condition.
%
The SMT extended version was not able to scale it to the desired results and either timed-out (T/O) or ran out of memory (M/O).
%
This can be due to the fact that directly translating a quantified formula into a quantifier-free by introducing a vector of variables for each boolean possibility can lead to exponential increase in the formula size.
%
%
\subsubsection{$\fqbf$ search encoding of the universal condition}
We will describe several constraints that encode universal condition based on QBF formulation in this section.
%
To avoid cumbersome notation, we will not explicitly write the ranges
of the indexing in the constraints.
%
$i$ and $j$ will range over nodes, i.e., from $1$ to $\nu$.
%
$m$ will range over molecules, i.e., from $1$ to $\mu$.
%
$q$ will range over edges between two nodes, i.e., from $1$ to $\pi$.
%

@@

Now we will consider the encodings for the
listed synthesis problems.
%
The presented variations represent the encodings
supported by our tool.
%
Additionally, the combinations of the variation are also possible and
our tool easily supports them.
%
For simplicity of the presentation, we assume that if we are
synthesizing an aspect of VTS, then all other aspects are fully given.
%
Therefore, we will describe two kinds of constraints for synthesis
problems.
%
One will encode the variable part in the synthesis problem and
the other encodes the fixed parts.
%
Subsequently, the two constraints will be put together with 
$\texttt{Consistancy}$ and 
$\texttt{Property}$ to construct the constraints for synthesis.

\subsubsection{Fixing VTS by adding edges}
%
Now we will consider the case when we add new edges to VTS to satisfy the properties.
%
In the following, the pseudo-Boolean formula $\texttt{AddE}$ encodes
that at most $slimit$ new undeclared edges may be added in the VTS.
%
\texttt{FixedForEdge} encodes the parts of the VTS that are not allowed to change.
\begin{align*}
&\texttt{AddE}(slimit) = \sum~\texttt{AbsentE} \leq slimit\\
&\texttt{FixedForEdge} = \texttt{PresentCons} \land \texttt{UnknownActive} \land\\
& \quad\quad\quad\quad \lnot \Lor \texttt{AbsentELabel} \;\union\;
\texttt{AbsentNLabel} \;\union\;
\texttt{AbsentP}
\end{align*}
We put together the constraints and obtain the following formula.
\begin{align*}
& \texttt{SynthE}(slimit) =
\texttt{Consistancy}\land \texttt{Property} \land
\texttt{FixedForEdge} \land
\texttt{AddE}(slimit)
% \tag{SynthE}\label{eq:addedge1}
\end{align*}
Similar to what we have seen $\texttt{Consistancy}$ encodes the basic constraints about VTS,
$\texttt{Property}$\; encodes the goal, and
the rest two are defined just above.
%
A satisfying model of $\texttt{SynthE}$ will make 
some of the edges in $\texttt{AbsentE}$ true such that~$\texttt{Property}$ is satisfied.
%
We limit the addition of the edges since we look for a fix that requires a minimum number
of changes in the given VTS.
%
We start with $slimit = 1$ and grow one by one until $\texttt{SynthE}(slimit)$
becomes satisfiable.

In the later synthesis problems, we will construct a similar QBF
formula with same first two parts and the last two are due the
requirements of the synthesis problem.
%
\subsubsection{Fixing VTS by adding molecules to the labels}
The system may also be fixed only by modifying labels on the edges or the nodes instead
of adding edges.
%
Here let us consider only adding molecules to the labels of edges.
%
In the following, the formula encodes that only $slimit$ edge labels may be added.
%
\begin{align*}
&\texttt{AddLabelEdge}(slimit) = 
\sum~\texttt{UnknwonEdgeLabel}  \leq slimit\\
& \texttt{FixedForLabel} = \texttt{PresentCons} \land \texttt{UnknownActive} \land\\
& \quad \quad \lnot \Lor \texttt{AbsentE} \;\union\;
\texttt{AbsentNLabel} \;\union\;
\texttt{AbsentP}\\
&\texttt{SynthLabel}(slimit) = \\
& \quad\quad
(  \texttt{Consistancy}\land \texttt{Property} \land
\texttt{FixedForLabel} \land \texttt{AddLabelEdge}(slimit) )
\end{align*} 
%
Similar to the previous encoding, we solve the
satisfiability of the above formula to obtain additional molecules
that may be added to the edge labels to satisfy the properties.
%
\subsubsection{Fixing VTS by learning activity functions}
Now we consider a scenario where some of the activity functions for
some of the molecules are missing.
%
The activity functions are $\mu$-input Boolean functions.
%
First, we choose a class of formulas for the candidate functions.
%
We encode the candidates in a formula with parameters.
%
By assigning different values for the parameters, a solver may select
different candidates for the activity functions.
%
We will illustrate only one class of formulas.
%
However, we support other classes of formulas, for example, $k$-CNF.

@@

\subsubsection{Fixing VTS by both adding/deleting parts}
%
Now we will consider repairing of VTS by allowing not only addition but also
deletion of the molecules, edges, functions, or pairing matrix.
%
We have encoded the repairing in our tool by introducing flip bits
for each variable that is modifiable in the VTS.
%
We illustrate the repairing on one class of variables and rest can be
easily extended.
%
Let us consider repairing of node labels.
%
For each bit $n_{i,m}$, we create a bit $flip_{i,m}$.
%
We add constraints that take xor of VTS assigned values for  $n_{i,m}$
and $flip_{i,m}$.
%
We also limit the number of $flip_{i,m}$ that can be true, therefore
limiting the number of flips.
%
The above constraints are encoded in $\texttt{FlipN}(slimit)$.
\begin{align*}
&\texttt{FlipN}(slimit,flip) = \Land \{n_{i,m}\lxor flip_{i,m}| m \in \nlabel(i) \land i \in \nodes \} \land \\
& \quad \quad \quad
\Land \{ \lnot n_{i,m}\lxor flip_{i,m}| m \not \in \nlabel(i) \land i \in \nodes \} \land  \sum\limits_{i,m} flip_{i,m} \leq slimit
\end{align*}
%
Similar to the earlier variations, we construct
$\texttt{SynthRepairNode}(slimit)$ for the repair.
%
Where, the constraint $\texttt{FixedForNodeRepair}$ encodes all the parts of VTS that do not change.
\begin{align*}
% \\ 
& \texttt{FixedForNodeRepair} =  \texttt{PresentE} \land  
\texttt{PresentP} \land \texttt{KnownActiveE} 
\land \\
& \quad \texttt{UnknownActive} \land \texttt{KnownActiveN} \land \lnot 
\Lor \texttt{AbsentE} \;\union\; \texttt{AbsentNLabel} \;\union\;
\texttt{AbsentP}\\
&\texttt{SynthRepairNode}(slimit,flip) = \\
& \quad\quad
(  \texttt{Consistancy}\land \texttt{Property} \land
\texttt{FixedForNodeRepair} \land \texttt{FlipN}(slimit,flip) )
\end{align*}
A satisfying model of $\texttt{SynthRepairNode}(slimit,flip)$ will assign some
$flip$ bits to true.
We will learn from the assignments the needed modifications in the VTS.

>>>>

%if all the pairs of the present molecule are present on the same edge than all of them inhibit each other resulting in making all the molecules inactive, we term this process ``pairing inhibition".
%
The constraints for the variants can be given to a SAT or a SMT solver to find
VTSs that belong to the variants.
\subsection{Encoding for $\fqbf$ search problem}\label{enc:univ-cond}
%
\input{table2}
\noindent The specification for universal condition Section~\ref{sec:problem} requires quantifier alternation. 
%
Most SAT solvers can not efficiently handle the quantifier in the formulas. 
% 
Although, it is possible to encode the universal condition as a SAT or SMT problem.
%
There is always a danger of a blow-up in the formula size, making it harder for the SAT solver to handle. 
%Which may blow up the formula size for the solver to solve.
% 
As an experiment, we have implemented a direct SMT version (encoding not presented) of the universal problem extending the current $\fsmt$ search encodings.
%
The basic idea is to conjunct the formula for each combination value of the quantified variable. 
%
The Table~\ref{tab:satqbf-graph}, shows the search results of this encoding for the universal condition with 4 nodes and 3-connected graphs. 
%
The connectivity constraint and connectedness constraints are tagged with subscript $U$, $\texttt{StabilityC}_{U}$, $\texttt{Connected}(k)_{U}$,  to distinguish from earlier presented encodings of existential condition.
%
The SMT extended version was not able to scale it to the desired results and either timed-out (T/O) or ran out of memory (M/O).
%
This can be due to the fact that directly translating a quantified formula into a quantifier-free by introducing a vector of variables for each boolean possibility can lead to exponential increase in the formula size.
%
%
\subsubsection{$\fqbf$ search encoding of the universal condition}
We will describe several constraints that encode universal condition based on QBF formulation in this section.
%
To avoid cumbersome notation, we will not explicitly write the ranges
of the indexing in the constraints.
%
$i$ and $j$ will range over nodes, i.e., from $1$ to $\nu$.
%
$m$ will range over molecules, i.e., from $1$ to $\mu$.
%
$q$ will range over edges between two nodes, i.e., from $1$ to $\pi$.
%

@@

Now we will consider the encodings for the
listed synthesis problems.
%
The presented variations represent the encodings
supported by our tool.
%
Additionally, the combinations of the variation are also possible and
our tool easily supports them.
%
For simplicity of the presentation, we assume that if we are
synthesizing an aspect of VTS, then all other aspects are fully given.
%
Therefore, we will describe two kinds of constraints for synthesis
problems.
%
One will encode the variable part in the synthesis problem and
the other encodes the fixed components.
%
Subsequently, the two constraints will be put together with 
$\texttt{Consistancy}$ and 
$\texttt{Property}$ to construct the constraints for synthesis.

\subsubsection{Fixing VTS by adding edges}
%
Now we will consider the case when we add new edges to VTS to satisfy the properties.
%
In the following, the pseudo-Boolean formula $\texttt{AddE}$ encodes
that at most $slimit$ new undeclared edges may be added in the VTS.
%
\texttt{FixedForEdge} encodes the components of the VTS that are not allowed to change.
\begin{align*}
&\texttt{AddE}(slimit) = \sum~\texttt{AbsentE} \leq slimit\\
&\texttt{FixedForEdge} = \texttt{PresentCons} \land \texttt{UnknownActive} \land\\
& \quad\quad\quad\quad \lnot \Lor \texttt{AbsentELabel} \;\union\;
\texttt{AbsentNLabel} \;\union\;
\texttt{AbsentP}
\end{align*}
We put together the constraints and obtain the following formula.
\begin{align*}
& \texttt{SynthE}(slimit) =
\texttt{Consistancy}\land \texttt{Property} \land
\texttt{FixedForEdge} \land
\texttt{AddE}(slimit)
% \tag{SynthE}\label{eq:addedge1}
\end{align*}
Similar to what we have seen $\texttt{Consistancy}$ encodes the basic constraints about VTS,
$\texttt{Property}$\; encodes the goal, and
the rest two are defined just above.
%
A satisfying model of $\texttt{SynthE}$ will make 
some of the edges in $\texttt{AbsentE}$ true such that~$\texttt{Property}$ is satisfied.
%
We limit the addition of the edges since we look for a fix that requires a minimum number
of changes in the given VTS.
%
We start with $slimit = 1$ and grow one by one until $\texttt{SynthE}(slimit)$
becomes satisfiable.

In the later synthesis problems, we will construct a similar QBF
formula with same first two components and the last two are due the
requirements of the synthesis problem.
%
\subsubsection{Fixing VTS by adding molecules to the labels}
The system may also be fixed only by modifying labels on the edges or the nodes instead
of adding edges.
%
Here let us consider only adding molecules to the labels of edges.
%
In the following, the formula encodes that only $slimit$ edge labels may be added.
%
\begin{align*}
&\texttt{AddLabelEdge}(slimit) = 
\sum~\texttt{UnknwonEdgeLabel}  \leq slimit\\
& \texttt{FixedForLabel} = \texttt{PresentCons} \land \texttt{UnknownActive} \land\\
& \quad \quad \lnot \Lor \texttt{AbsentE} \;\union\;
\texttt{AbsentNLabel} \;\union\;
\texttt{AbsentP}\\
&\texttt{SynthLabel}(slimit) = \\
& \quad\quad
(  \texttt{Consistancy}\land \texttt{Property} \land
\texttt{FixedForLabel} \land \texttt{AddLabelEdge}(slimit) )
\end{align*} 
%
Similar to the previous encoding, we solve the
satisfiability of the above formula to obtain additional molecules
that may be added to the edge labels to satisfy the properties.
%
\subsubsection{Fixing VTS by learning activity functions}
Now we consider a scenario where some of the activity functions for
some of the molecules are missing.
%
The activity functions are $\mu$-input Boolean functions.
%
First, we choose a class of formulas for the candidate functions.
%
We encode the candidates in a formula with parameters.
%
By assigning different values for the parameters, a solver may select
different candidates for the activity functions.
%
We will illustrate only one class of formulas.
%
However, we support other classes of formulas, for example, $k$-CNF.

@@

\subsubsection{Fixing VTS by both adding/deleting components}
%
Now we will consider repairing of VTS by allowing not only addition but also
deletion of the molecules, edges, functions, or pairing matrix.
%
We have encoded the repairing in our tool by introducing flip bits
for each variable that is modifiable in the VTS.
%
We illustrate the repairing on one class of variables and rest can be
easily extended.
%
Let us consider repairing of node labels.
%
For each bit $n_{i,m}$, we create a bit $flip_{i,m}$.
%
We add constraints that take xor of VTS assigned values for  $n_{i,m}$
and $flip_{i,m}$.
%
We also limit the number of $flip_{i,m}$ that can be true, therefore
limiting the number of flips.
%
The above constraints are encoded in $\texttt{FlipN}(slimit)$.
\begin{align*}
&\texttt{FlipN}(slimit,flip) = \Land \{n_{i,m}\lxor flip_{i,m}| m \in \nlabel(i) \land i \in \nodes \} \land \\
& \quad \quad \quad
\Land \{ \lnot n_{i,m}\lxor flip_{i,m}| m \not \in \nlabel(i) \land i \in \nodes \} \land  \sum\limits_{i,m} flip_{i,m} \leq slimit
\end{align*}
%
Similar to the earlier variations, we construct
$\texttt{SynthRepairNode}(slimit)$ for the repair.
%
Where, the constraint $\texttt{FixedForNodeRepair}$ encodes all the components of VTS that do not change.
\begin{align*}
% \\ 
& \texttt{FixedForNodeRepair} =  \texttt{PresentE} \land  
\texttt{PresentP} \land \texttt{KnownActiveE} 
\land \\
& \quad \texttt{UnknownActive} \land \texttt{KnownActiveN} \land \lnot 
\Lor \texttt{AbsentE} \;\union\; \texttt{AbsentNLabel} \;\union\;
\texttt{AbsentP}\\
&\texttt{SynthRepairNode}(slimit,flip) = \\
& \quad\quad
(  \texttt{Consistancy}\land \texttt{Property} \land
\texttt{FixedForNodeRepair} \land \texttt{FlipN}(slimit,flip) )
\end{align*}
A satisfying model of $\texttt{SynthRepairNode}(slimit,flip)$ will assign some
$flip$ bits to true.
We will learn from the assignments the needed modifications in the VTS.
